---
title: "Homework for 31 Jan"
author: "Catherine Galley"
date: "27/01/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5.2.4 Exercises
```{r}
library(nycflights13)
library(tidyverse)
d <- nycflights13::flights
head(d)
```


Find all flights that


Had an arrival delay of two or more hours
```{r}


```
Flew to Houston (IAH or HOU)
Were operated by United, American, or Delta
Departed in summer (July, August, and September)
Arrived more than two hours late, but didn’t leave late
Were delayed by at least an hour, but made up over 30 minutes in flight
Departed between midnight and 6am (inclusive)
Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)