---
title: "Homework for 31 Jan"
author: "Catherine Galley"
date: "27/01/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5.2.4 Exercises
```{r}
library(nycflights13)
library(tidyverse)
d <- nycflights13::flights
head(d)
```


Find all flights that


Had an arrival delay of two or more hours
```{r}
summary(d$arr_delay)
filter(d, arr_delay >= 120)

```

Flew to Houston (IAH or HOU)
```{r}
filter(d, dest == "IAH" | dest == "HOU")
```

Were operated by United, American, or Delta
```{r}
filter(d, carrier == "UA" | carrier == "AA" | carrier == "DL")
```
Departed in summer (July, August, and September)
```{r}
filter(d, month %in% c(7, 8, 9))
 
```
Arrived more than two hours late, but didn’t leave late
```{r}
filter(d, arr_delay > 120 & dep_delay <= 0)
```
Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(d, dep_delay >= 60 & arr_delay <= 30)
```
Departed between midnight and 6am (inclusive)
```{r}
filter(d, dep_time %in% c(0000, 0600))
```
Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}

```

How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
length(is.na(d$dep_time))
summary(d$dep_time)
summary(d)
```
8255 Missing Values = Cancelled Flights.
Also missing = dep_delay, arr_time, arr_delay, air_time.

Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

# 5.5.2 Exercises

Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

What does 1:3 + 1:10 return? Why?

What trigonometric functions does R provide?


