---
title: "Homework for Week 4"
author: "Catherine Galley"
date: "02/02/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(nycflights13)
flights<- nycflights13::flights
airports <- nycflights13::airports
airlines  <- nycflights13::airlines
planes <- nycflights13::planes
weather <- nycflights13::weather


```

## 13.3.1 Exercises

Add a surrogate key to flights.

```{r}
flights <- flights %>%
  arrange(year, month, day, sched_dep_time, carrier, flight) %>%
  mutate(flight_id = row_number()) %>%
  glimpse()
head(flights)
```

Identify the keys in the following datasets

Lahman::Batting,
```{r}
install.packages("Lahman")
library(Lahman)
batting <- Lahman::Batting
colnames(batting)
batting %>%
  count(playerID, yearID, stint) %>%
  filter(n>1)
```
babynames::babynames
```{r}
install.packages("babynames")
library(babynames)
bn <- babynames::babynames
colnames(bn)
bn %>%
  count(sex, year, name) %>%
  filter(nn > 1)
```
nasaweather::atmos
```{r}
install.packages("nasaweather")
library(nasaweather)
atmos <- nasaweather::atmos
colnames(atmos)
atmos %>%
  count(lat, long, year, month) %>%
  filter (n > 1)
```
fueleconomy::vehicles
```{r}
install.packages("fueleconomy")
library(fueleconomy)
vehicles <- fueleconomy::vehicles
colnames(vehicles)
vehicles %>%
  count(id) %>%
  filter(n > 1)

```
ggplot2::diamonds
```{r}
diamonds <- ggplot2::diamonds
colnames(diamonds)
diamonds %>%
  count(x, y, z, cut, clarity, carat, color, depth, table, price) %>%
  filter(n > 1)
```
(You might need to install some packages and read some documentation.)

Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers.
```{r}
master <- Lahman::Master
salaries <- Lahman::Salaries
colnames(batting)
colnames(master)
colnames(salaries)

managers <- Lahman::Managers
am <- Lahman::AwardsManagers
colnames(managers)
colnames(am)
```
Batting and Master: Player ID
Master and Salaries: Player ID
Salaries and BattingL Player ID, Year ID, LgID, Team ID

Master and Managers: playerID
Master and AwardsManagers: playerID
Managers and AwardsManagers: playerID, yearID, lgID

How would you characterise the relationship between the Batting, Pitching, and Fielding tables?
```{r}

```


## Mutating Joins
Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:
```{r}
library(maps)
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)

flights2 <- nycflights13::flights

colnames(airports)
colnames(flights)
flights <- flights %>%
  group_by(origin, dest) %>%
  filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
  mutate(total_delay = arr_delay - dep_delay) %>%
  summarise(
    mean_total_delay = mean(total_delay)
  )
summary(flights$mean_total_delay)
flights_grouped <- flights %>% 
  mutate(delay_category = case_when(
  between(mean_total_delay, -30, -20) ~ "20",
  between(mean_total_delay, -19.9999, -10) ~ "-10",
  between(mean_total_delay, -9.9999, 1) ~ "-1",
  between(mean_total_delay, -0.9999, 0.9999) ~ "0",
  between(mean_total_delay, 1, 10.9999) ~ "1",
  between(mean_total_delay, 11, 20) ~ "10"
)) %>%
  inner_join(airports, by = c(dest = "faa"))

      
flights_grouped %>%
  ggplot(aes(lon, lat, colour = mean_total_delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap()
``` 


You might want to use the size or colour of the points to display the average delay for each airport.


Add the location of the origin and destination (i.e. the lat and lon) to flights.
```{r}
flights2 <- nycflights13::flights

airport_locations <- airports %>%
  select(faa, lat, lon)

flights2 %>%
  select(year:day, hour, origin, dest) %>%
  left_join(
    airport_locations,
    by = c("origin" = "faa")
  ) %>%
  left_join(
    airport_locations,
    by = c("dest" = "faa"))

``` 
Is there a relationship between the age of a plane and its delays?

```{r}
colnames(planes)
colnames(flights)
planes %>%
  count(tailnum) %>%
  filter(n > 1)

flights %>%
  filter(!is.na(dep_time))
  count(tailnum, year, month, day, origin, dep_delay, arr_delay, dest, dep_time) %>%
  filter(n > 1)

plane_age <- planes %>%
  select(tailnum, year)

flight_plane <- flights %>%
  select(year, month, day, tailnum, origin, dep_delay, arr_delay, dest, dep_time) %>%
  inner_join(planes, by = tailnum, "tailnum")
```
```{r}
plane_cohorts <- inner_join(flights,
  select(planes, tailnum, plane_year = year),
  by = "tailnum"
) %>%
  mutate(age = year - plane_year) %>%
  filter(!is.na(age)) %>%
  group_by(age) %>%
  summarise(
    dep_delay_mean = mean(dep_delay, na.rm = TRUE),
    dep_delay_sd = sd(dep_delay, na.rm = TRUE),
    arr_delay_mean = mean(arr_delay, na.rm = TRUE),
    arr_delay_sd = sd(arr_delay, na.rm = TRUE),
    n = n()
  )
  
ggplot(plane_cohorts, aes(x = age, y = dep_delay_mean)) +
  geom_ribbon(
    mapping = aes(
      ymin = dep_delay_mean -
        2 * dep_delay_sd / sqrt(n),
      ymax = dep_delay_mean +
        2 * dep_delay_sd / sqrt(n)
    ),
    alpha = 0.3
  ) +
  geom_point() +
  scale_x_continuous("Age of plane (years)", breaks = seq(0, 50, by = 10)) +
  scale_y_continuous("Mean Departure Delay (minutes)")
``` 
```{r}
ggplot(plane_cohorts, aes(x = age, y = arr_delay_mean)) +
  geom_ribbon(
    mapping = aes(
      ymin = arr_delay_mean -
        2 * arr_delay_sd / sqrt(n),
      ymax = arr_delay_mean +
        2 * arr_delay_sd / sqrt(n)
    ),
    alpha = 0.3
  ) +
  geom_point() +
  scale_x_continuous("Age of Plane (years)", breaks = seq(0, 50, by = 10)) +
  scale_y_continuous("Mean Arrival Delay (minutes)")
```
What weather conditions make it more likely to see a delay?
```{r}
colnames(weather)
delay_weather <- weather %>%
  select(origin, year, month, day, temp, wind_speed, wind_gust, precip, pressure, hour) %>%
  inner_join(flights, 
             select(year, month, day, dep_delay, arr_delay, origin, hour, dest),
             by = c("origin" = "origin",
                    "day" = "day",
                    "year" = "year",
                    "month" = "month",
                    "hour" = "hour"))
head(delay_weather)

delays <- delay_weather %>%
  filter(!is.na(dep_delay), !is.na(arr_delay)) %>%
  group_by(origin, temp, wind_speed, wind_gust, pressure, precip) %>%
  summarise(
    mean_dep_delay = mean(dep_delay),
    mean_arr_delay = mean(arr_delay))
  
summary(delays$mean_arr_delay)
summary(delays$mean_dep_delay)

plot(delays$mean_dep_delay, delays$temp)
plot(delays$mean_dep_delay, delays$wind_speed)
plot(delays$mean_dep_delay, delays$wind_gust)
plot(delays$mean_dep_delay, delays$precip)
plot(delays$mean_dep_delay, delays$pressure)
```

What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.
```{r}
install.packages("viridis")
library(viridis)


flights %>% filter( 
       day == 13,
       month == 6,
       year == 2013) %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(y = lat, x = lon, size = delay, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap() +
  scale_colour_viridis()

```

## Reading Data into R - Waves 1 and 2
```{r}
rm(list = ls())
library(data.table)
UndSoc1 <- fread("/Users/catherinegalley/Documents/DataThree/data/UKDA-6614-tab/tab/ukhls_w1/a_indresp.tab")
UndSoc2 <- fread("/Users/catherinegalley/Documents/DataThree/data/UKDA-6614-tab/tab/ukhls_w2/b_indresp.tab")

UndSoc1ed <- fread("/Users/catherinegalley/Documents/DataThree/data/UKDA-6614-tab/tab/ukhls_w1/a_indresp.tab",
                   select = c("pidp", "a_sex", "a_dvage", "a_vote6"))

UndSoc2ed <- fread("/Users/catherinegalley/Documents/DataThree/data/UKDA-6614-tab/tab/ukhls_w2/b_indresp.tab",
                   select = c("pidp", "b_sex", "b_dvage", "b_vote6"))
rm(UndSoc1, UndSoc2)
```

## Joining the Two Waves 
```{r}

inner <- UndSoc1ed %>%
        inner_join(UndSoc2ed, by = "pidp")

left <- UndSoc1ed %>%
        left_join(UndSoc2ed, by = "pidp")

right <- UndSoc1ed %>%
        right_join(UndSoc2ed, by = "pidp")

full <- UndSoc1ed %>%
        full_join(UndSoc2ed, by = "pidp")
```
