---
title: "Homework for 31 Jan"
author: "Catherine Galley"
date: "27/01/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 5.2.4 Exercises
```{r}
library(nycflights13)
library(tidyverse)
d <- nycflights13::flights
flights <- nycflights13::flights
head(d)
head(d, n = 30)
```


Find all flights that


Had an arrival delay of two or more hours
```{r}
summary(d$arr_delay)
filter(d, arr_delay >= 120)

```

Flew to Houston (IAH or HOU)
```{r}
filter(d, dest == "IAH" | dest == "HOU")
```

Were operated by United, American, or Delta
```{r}
filter(d, carrier == "UA" | carrier == "AA" | carrier == "DL")
```
Departed in summer (July, August, and September)
```{r}
filter(d, month %in% c(7, 8, 9))
 
```
Arrived more than two hours late, but didn’t leave late
```{r}
filter(d, arr_delay > 120 & dep_delay <= 0)
```
Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(d, dep_delay >= 60 & arr_delay <= 30)
```
Departed between midnight and 6am (inclusive)
```{r}
filter(d, dep_time %in% c(0000, 0600))
```
Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}

```

How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
length(is.na(d$dep_time))
summary(d$dep_time)
summary(d)
```
8255 Missing Values = Cancelled Flights.
Also missing = dep_delay, arr_time, arr_delay, air_time.

Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

# Mutate - Practice from Chapter
```{r}
flights_sml <- select(d, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)

mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)

transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

# 5.5.2 Exercises

Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
summary(d$dep_time)
flights_times <- mutate(flights,
  dep_time_mins = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
  sched_dep_time_mins = (sched_dep_time %/% 100 * 60 +
    sched_dep_time %% 100) %% 1440)
select(
  flights_times, dep_time, dep_time_mins, sched_dep_time, sched_dep_time_mins
)
```

Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
```{r}
flights_airtime <- 
  mutate(flights,
    dep_time_minutes = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    arr_time_minutes = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440,
    air_time_diff = air_time - arr_time_minutes + dep_time_minutes
)
nrow(filter(flights_airtime, air_time_diff != 0))

ggplot(flights_airtime, aes(x = air_time_diff)) + geom_histogram(binwidth = 1)
```

Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

```{r}
flights_deptime <- 
  mutate(flights,
    dep_time_min = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    sched_dep_time_min = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% 1440,
    dep_delay_diff = dep_delay - dep_time_min + sched_dep_time_min
)
filter(flights_deptime, dep_delay_diff != 0)

ggplot(
  filter(flights_deptime, dep_delay_diff > 0),
  aes(y = sched_dep_time_min, x = dep_delay_diff)
) + 
  geom_point()
```

Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r}
flights_delayed <- mutate(flights, dep_delay_rank = min_rank(-dep_delay))
flights_delayed <- filter(flights_delayed, dep_delay_rank <= 20)
arrange(flights_delayed, dep_delay_rank)
```

What does 1:3 + 1:10 return? Why?
```{r}
1:3 + 1:10

```

What trigonometric functions does R provide?
Cos, Sin, Tan, 
ArcCos, ArcSin, ArcTan

## Chapter 3 - Website

```{r}
library(data.table)
library(tidyverse)
W1 <- read_tsv("/Users/catherinegalley/Documents/DataThree/data/UKDA-6614-tab/tab/ukhls_w1/a_indresp.tab")

table(W1$a_vote6)
prop.table(table(W1$a_vote6))
library(knitr)
kable(data.table(prop.table(table(W1$a_vote6))),digits=2)

W1$a_vote6 %>%
  table() %>%
  prop.table() %>%
  data.frame() %>%
  kable(digits = 2) 
  
```
# Sumarise

```{r}
by_day <- group_by(d, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))

delays <- d %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")

ggplot(data = delays, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )

ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```

# 5.6.7 Exercises 

Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.

A flight is always 10 minutes late.

A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.

99% of the time a flight is on time. 1% of the time it’s 2 hours late.

Which is more important: arrival delay or departure delay?

Arrival Delay - most likely to affect onward travel
Consistency less of a problem than variation 

Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).

```{r}

not_cancelled <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>%
  group_by(dest) %>%
  summarise(n = length(dest))

not_cancelled %>%
  group_by(dest) %>%
  summarise(n = n())

```

Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?

Most important column is arr_delay. 
```{r}

filter(flights, !is.na(dep_delay), is.na(arr_delay)) %>%
  select(dep_time, arr_time, sched_arr_time, dep_delay, arr_delay)
```

Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}

cancelled_delayed <- 
  flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(
    prop_cancelled = mean(cancelled),
    avg_dep_delay = mean(dep_delay, na.rm = T)
)

ggplot(cancelled_delayed, aes(x = avg_dep_delay, prop_cancelled)) + geom_point() + geom_smooth()
``` 
Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))

What does the sort argument to count() do. When might you use it?


