---
title: "Homework - Tidy Data"
author: "Catherine Galley"
date: "08/02/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## 12.2.1

Using prose, describe how the variables and observations are organised in each of the sample tables.

Table 1: Each variable has its own column and each observation its own row

Table 2: Cases and Population share a column

Table 3: Cases and Population have been combined to make rate, with the original data much harder to view

Table 4: Two seperate tables, one for cases and one for population = harder to compare

Compute the rate for table2, and table4a + table4b. You will need to perform four operations:


Extract the number of TB cases per country per year.
Extract the matching population per country per year.
Divide cases by population, and multiply by 10000.
Store back in the appropriate place.
Which representation is easiest to work with? Which is hardest? Why?

Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?

## Gather
```{r}
tidy4a <- table4a %>% gather(`1999`, `2000`, key = "year", value = "cases")

tidy4b <- table4b %>% gather(`1999`, `2000`, key = "year", value = "population")

left_join(tidy4a, tidy4b)
```
## Spread
```{r}
table2 %>%
  spread(key = type, value = count)
```

# Ex 12.3.3 
Not symetrical because with gather, you have to specify the columns you want to select, the key and they new variable you want to make, but with spread you have to specify the columns you want and they will be seperated

2. because the years are not in `` so they are recognised as numbers not variables. 

3. Because PW has two values for the same key. could add a unique identifier for each variable

4. gather by sex and pregnant and then count




## Separate

```{r}
table3 %>%
  separate(rate, into = c("cases", "population"), sep = "/", convert = T)
```
# Unite 
```{r}
table5 %>% unite(new, century, year, sep = "")
```
## 12.4.3

What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
  
  The extra argument tells separate() what to do if there are too many pieces, and the fill argument tells it what to do if there aren’t enough. By default, separate() drops extra values with a warning.
  
  
Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?

The remove argument discards input columns in the result data frame. You would set it to FALSE if you want to create a new variable, but keep the old one

Compare and contrast separate() and extract(). Why are there three variations of separation (by position, by separator, and with groups), but only one unite?

Seperate = splits a column into multiple columns by separator

Extract = uses a regular expression to specify groups in character vector and split that single character vector into multiple columns

Seperate and Extract = one column to many 
Unite = many columns to one 
# WHO dataset 
```{r}
who <- tidyr::who
who1 <- who %>%
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = T)

who1 %>%
  count(key)

who2 <- who1 %>%
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))

who3 <- who2 %>%
  separate(key, c("new", "type", "sexage"), sep = "_")

who4 <- who3 %>%
  select(-new, -iso2, -iso3)

who5 <- who4 %>%
  separate(sexage, c("sex", "age"), sep = 1)
```