---
title: "Lecture 6"
author: "Catherine Galley"
date: "21/02/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(cache = TRUE)
```
Plan for today:

- Assignment 2: solution
- Assignment 3
- Simple data visualisation 
- Homework for next week

## Exercises

1. Read the *indresp* file from Wave 8 and keep the following variables: pidp, derived sex and age, ethnic group (h_indresp), government office region (h_gor_dv), and net personal income (h_fimnnet_dv).
```{r}
library(tidyverse)
d <- read_tsv("/Users/catherinegalley/Documents/DataThree/data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
d <- d %>%
  select(pidp, h_sex_dv, h_age_dv, h_ethn_dv, h_gor_dv, h_fimnnet_dv)

d$h_sex_dv[d$h_sex_dv == 0] <- NA
d$h_gor_dv[d$h_gor_dv == -9] <- NA
```
For all charts use ggplot2. You may need to clean and recode variable before visualising.

We will start with univariate distributions.

2. Visualise the distribution of income with a histogram, a density plot and a box plot.
```{r}
ggplot(d, aes(h_fimnnet_dv)) + geom_histogram(binwidth = 50) + xlim(c(-100, 15000))
ggplot(d, aes(h_fimnnet_dv)) + geom_density() + xlim(c(-100, 5000))
ggplot(d, aes(y = h_fimnnet_dv)) + geom_boxplot() + ylim(1, 5000)
?geom_boxplot
  
```
3. Visualise the distribution of sex with a bar graph. 
```{r}
ggplot(d, aes(h_sex_dv)) + geom_bar(fill = "navyblue")
```
Bivariate distributions.

4. Create a bar chart showing mean income by region. Then make a grouped bar chart showing mean income by region for men and women separately.
```{r}
d %>%
  mutate(region = case_when(
    h_gor_dv == 1 ~ "North East",
    h_gor_dv == 2 ~ "North West",
    h_gor_dv == 3 ~ "Yorkshire and the Humber",
    h_gor_dv == 4 ~ "East Midlands",
    h_gor_dv == 5 ~ "West Midlands",
    h_gor_dv == 6 ~ "East of England",
    h_gor_dv == 7 ~ "London",
    h_gor_dv == 8 ~ "South East",
    h_gor_dv == 9 ~ "South West",
    h_gor_dv == 10 ~ "Wales",
    h_gor_dv == 11 ~ "Scotland",
    h_gor_dv == 12 ~ "Northern Ireland"
  )) %>%
  group_by(region) %>%
  summarise(median_income = median(h_fimnnet_dv))  %>%
  ggplot(aes(x = region, y = median_income)) +
  geom_bar(stat = "identity", fill = "turquoise2") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

d %>%
  mutate(region = case_when(
    h_gor_dv == 1 ~ "North East",
    h_gor_dv == 2 ~ "North West",
    h_gor_dv == 3 ~ "Yorkshire and the Humber",
    h_gor_dv == 4 ~ "East Midlands",
    h_gor_dv == 5 ~ "West Midlands",
    h_gor_dv == 6 ~ "East of England",
    h_gor_dv == 7 ~ "London",
    h_gor_dv == 8 ~ "South East",
    h_gor_dv == 9 ~ "South West",
    h_gor_dv == 10 ~ "Wales",
    h_gor_dv == 11 ~ "Scotland",
    h_gor_dv == 12 ~ "Northern Ireland"
  )) %>%
  filter(!is.na(h_sex_dv)) %>%
  filter(!is.na(region)) %>%
  group_by(region, h_sex_dv) %>%
  summarise(median_income = median(h_fimnnet_dv))  %>%
  ggplot(aes(x = region, y = median_income, fill = h_sex_dv)) +
  geom_bar(stat = "identity", position = "dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

#ggplot(d2, aes(x = h_gor_dv, y = mean_income)) + geom_bar(stat = "count")

#ggplot(d, aes(x = h_gor_dv, y = h_fimnnet_dv)) + geom_bar(stat = "identity", colour = "pink")

# ggplot(d, aes(x = h_gor_dv, y = h_fimnnet_dv, fill = as.factor(h_sex_dv))) + geom_bar(stat = "identity", position = "dodge")

```
5. Make a dot plot showing the same information as above (without splitting by sex). Sort regions in the descending order by mean income.
```{r}
d %>%
  mutate(region = case_when(
    h_gor_dv == 1 ~ "North East",
    h_gor_dv == 2 ~ "North West",
    h_gor_dv == 3 ~ "Yorkshire and the Humber",
    h_gor_dv == 4 ~ "East Midlands",
    h_gor_dv == 5 ~ "West Midlands",
    h_gor_dv == 6 ~ "East of England",
    h_gor_dv == 7 ~ "London",
    h_gor_dv == 8 ~ "South East",
    h_gor_dv == 9 ~ "South West",
    h_gor_dv == 10 ~ "Wales",
    h_gor_dv == 11 ~ "Scotland",
    h_gor_dv == 12 ~ "Northern Ireland"
  )) %>%
  group_by(region) %>%
  summarise(median_income = median(h_fimnnet_dv))  %>%
  ggplot(aes(x = region, y = median_income)) +
  geom_dotplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
6. Make a line chart showing mean income by age. Make another chart showing the association for men and women separately.
```{r}

```
7. Create a scatterplot showing the association between age and income for randomly selected 100 men in the sample. Now display a smoothing curve on the same graph showing the association between age and income.
```{r}

```